Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain

### 2018-10-31 pkg/disttree/inst/draft/draft.Rnw
changed numbering of supplements (A and B instead of 1 and 2)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2369 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 9093ae989a286d620c85bcfcb92b68ad87d3c504

\begin{itemize}
  \item \emph{Generalized linear models} (GLMs, \citealp{Nelder+Wedderburn:1972})
    encompassing an additional nonlinear link function for the conditional mean.
  \item \emph{Generalized additive models} (GAMs, \citealp{Hastie+Tibshirani:1986})
    allowing for smooth nonlinear effects in the covariates
    (Figure~\ref{fig:gam}, middle).
  \item \emph{Generalized additive models for location, scale, and shape}
    (GAMLSS, \citealt{Rigby+Stasinopoulos:2005}) adopting a probabilistic modeling
    approach. In GAMLSS, each parameter of a statistical distribution can depend
    on an additive predictor of the covariates comprising linear and/or
    smooth nonlinear terms (Figure~\ref{fig:gam}, right).
\end{itemize}
Thus, the above-mentioned models provide a broad toolbox for capturing different
aspects of the response (mean only vs.\ full distribution) and different types
of dependencies on the covariates (linear vs.\ nonlinear additive terms).

@@

In summary, the models discussed so far provide a broad and powerful toolset for
parametric distributional fits depending on a specified set of additive linear or
smooth nonlinear terms. A rather different approach to capturing the dependence on
covariates are tree-based models.
\begin{itemize}
  \item \emph{Regression trees} (\citealt{Breiman+Friedman+Stone:1984}) recursively
    split the data into more homogeneous subgroups and can thus capture abrupt shifts
    (Figure~\ref{fig:tree}, top left) and approximate nonlinear functions.
    Furthermore, trees automatically carry out a forward selection of covariates and
    their interactions. 
  \item \emph{Random forests} (\citealt{Breiman:2001}) average the predictions of
    an ensemble of trees fitted to resampled versions of the learning data. This
    stabilizes the recursive partitions from individual trees and hence better
    approximates smooth functions (Figure~\ref{fig:tree}, top middle).
\end{itemize}
While classical regression trees and random forests only model the mean of the response
we propose to follow the ideas from GAMLSS modeling -- as outlined
in Figure~\ref{fig:gam} -- and combine tree-based methods with parametric distributional models,
yielding two novel techniques:
\begin{itemize}
  \item \emph{Distributional regression trees} (for short: \emph{distributional trees}) split the data into more homogeneous groups with
    respect to a parametric distribution, thus capturing changes in any distribution
    parameter like location, scale, or shape (Figure~\ref{fig:tree}, bottom left).
  \item \emph{Distributional regression forests} (for short: \emph{distributional forests}) utilize an ensemble of distributional trees
    for obtaining stabilized and smoothed parametric predictions (Figure~\ref{fig:tree},
    top right).
\end{itemize}

@@

Learning and validation data consist of observed 
daily precipitation sums provided by the National Hydrographical Service 
(\citealt{ehyd}) and numerical weather forecasts from the U.S.~National
Oceanic and Atmospheric Administration (NOAA).
Both, observations and forecasts are available for 1985--2012 and
the analysis is exemplified using July, the month with the most precipitation
in Tyrol.

@@

The predictive performance in terms of full probabilistic forecasts is
assessed using the continuous ranked probability score (CRPS, \citealt{Hersbach:2000}).
For each of the models this assesses the discrepancy of the
predicted distribution function $F$ from the observation $y$ by
$$
\text{CRPS}(y, F) = \int_{-\infty}^\infty (F(z) - \mathbf{1}(y \leq z))^2 dz
$$
where \(\mathbf{1}(\cdot)\) is the indicator function. In the subsequent
applications, the mean CRPS is always evaluated out of sample, 
either using cross-validation or a hold-out data set (2009--2012)
that was not used for learning (1985--2008). CRPS is a proper scoring rule
\citep{Gneiting+Raftery:2007} often used within the meteorological community.
Lower values indicate better performance.

@@

To assess calibration Figure~\ref{fig:axams-qq} shows residual QQ~plots for
out-of-sample predictions (2009--2012) from the different models trained on
1985--2008. Due to the point masses at zero 100 draws from the randomized
quantile residuals \citep{Dunn+Smyth:1996} are plotted in semi-transparent
gray. Overall, the randomized quantile residuals conform quite well with the
theoretical standard normal quantile (i.e., form a straight line close to the 
diagonal), indicating that all four models are sufficiently well calibrated. 
\hlc{This is also supported by the corresponding probability integral transform}
(PIT, \citealt{Gneiting+Balabdaoui+Raftery:2007}) \hlc{histograms for station
Axams in Supplement 2} \citep{Schlosser+Hothorn+Stauffer:2018b} \hlc{which 
contains a more detailed explanation of residual QQ plots and PIT histograms.}

@@

\subsubsection{Distributional specifications}
\hlc{Choosing an adequate distributional family is a crucial step and forms
the basis required to build a well fitting model. Therefore, this
decision should be made thoughtfully. 
In the presented application we applied a Gaussian distribution left-censored 
at zero as this has proven to be a well working choice in
earlier literature} \citep{Stauffer+Mayr+Messner:2017}. 
\hlc{To test for robustness against distributional misspecifation 
different distributional assumptions have been considered as well. 
In Supplement 1} \citep{Schlosser+Hothorn+Stauffer:2018a} \hlc{the exact
same evaluations as in} Section~\ref{sec:all} \hlc{are carried out for 15
observation stations, once employing a Gaussian hurdle model and once
with a logistic distribution left-censored at zero in order to account
for heavier tails. Under both assumptions the obtained results 
are similar to those presented in the paper applying a Gaussian distribution
left-censored at zero. In some settings the hurdle model leads to slight 
improvements, but this comes with more variability (especially for EMOS and the 
boosted GAMLSS). For the boosted GAMLSS the left-censored logistic distribution
shows slightly better results regarding CRPS skill score, however, only within 
a range of less than $2\%$ and not for all stations. 
Altogether, the distributional forest has proven to be 
very robust against distributional misspecifications.}

@@

\subsubsection*{East Tyrol}
Figure~\ref{fig:all-map} \hlc{shows that, contrary to the rest of Tyrol, 
for the majority of stations in East Tyrol the prespecified GAMLSS performs
best in this particular setting (learning data: 1985--2008, testing data:
2009--2012). East Tyrol is located south of the main Alpine Ridge and 
exhibits different climatological characteristics than the part north of 
the Alpine Ridge. It seems that the choice of covariates and their interactions 
we made for the prespecified GAMLSS method is particularly adequate for the 
region of East Tyrol for this specific analysis. However, the advantages of 
the prespecified GAMLSS models in this particular setting (training: 1985--2008, 
validation: 2009--2012) even out in a cross-validation framework for most of 
the stations in East Tyrol. These results are presented in Supplement 2} 
\citep{Schlosser+Hothorn+Stauffer:2018b} \hlc{where the exact same evaluations as in} 
Section~\ref{sec:axams} \hlc{for station Axams have been carried out 
for 14 other stations.}

\subsubsection*{Choice of station Axams}
%% choice of station Axams mit Verweis auf Supplement stationwise
\hlc{Axams was chosen for the more extensive evaluations presented
in this paper as it is the observation station being closest 
to Innsbruck, the capital of Tyrol and the work place of 
three of the authors, but not based on any other particular 
features. Furthermore, it has been checked, whether it 
is an adequate representative of all stations regarding 
the results of our evaluations. This is also confirmed by 
the extended analysis in Supplement 2} \citep{Schlosser+Hothorn+Stauffer:2018b}
\hlc{where the 15 considered stations also include some of those where 
the forest does not perform better than the other models in the setting 
presented in} Figure~\ref{fig:all-map}. \hlc{It shows that the differences in 
performance even out in the cross-validation framework where the forest 
performs at least equally well compared to the best of the other models 
at most stations.}

@@

The proposed methods are in the \textsf{R} package \textbf{disttree}
(version~\Sexpr{packageVersion("disttree")}) based on the \textbf{partykit}
package (version~\Sexpr{packageVersion("partykit")}), both available 
on \textsf{R}-Forge at
(\url{https://R-Forge.R-project.org/projects/partykit/}). The function
\code{distfit} fits distributional models by maximum likelihood, which is
used as the basis for the tree-building function \code{disttree}, upon which
the \code{distforest} is built. All functions can either be used with
GAMLSS family objects from the \textsf{R} package \textbf{gamlss.dist} 
(\citealt{Stasinopoulos+Rigby:2007}, version~\Sexpr{packageVersion("gamlss.dist")})
or with custom lists containing all required information about the distribution family.

@@

The fitted distributional forest for July 24 and
observation station Axams (including Figure~\ref{fig:axams-dist})
is reproducible using 
\code{demo("RainAxams", package = "disttree")}.
This also includes fitting the other zero-censored
Gaussian models considered in this paper and generating the corresponding 
QQ~plots (Figure~\ref{fig:axams-qq}) and PIT histograms 
\citep[Supplement 2]{Schlosser+Hothorn+Stauffer:2018b}.
Full replication of all results can be obtained with
\code{demo("RainTyrol", package = "disttree")}
requiring the companion \textsf{R} package \textbf{RainTyrol} (version~\Sexpr{packageVersion("RainTyrol")}),
also available within the \textsf{R}-Forge~project.
\hlc{The results presented in Supplement 1} \citep{Schlosser+Hothorn+Stauffer:2018a} and 
Supplement 2 \citep{Schlosser+Hothorn+Stauffer:2018b} \hlc{can be reproduced using}
\code{demo("RainDistributions", package = "disttree")} and 
\code{demo("RainStationwise", package = "disttree")} respectively.

@@

%\begin{supplement}[id=supp1]
%  \sname{Supplement 1}
%  \stitle{Different Response Distributions}
%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
%  \sdatatype{.pdf}" 
%  \sdescription{
%  In the case study presented in the main manuscript ``Distributional Regression
%  Forests for Probabilistic Precipitation Forecasting in Complex Terrain''
%  the novel distributional regression forests are based on a Gaussian response distribution,
%  left-censored at zero to accomodate both the non-negativity of precipitation and
%  the point mass at zero. To assess the goodness of fit of this response distribution
%  this supplement employs the same evaluations as in the main manuscript but based on
%  two other distributional assumptions: A logistic distribution, left-censored at zero,
%  is employed to potentially better capture heavy tails -- and a two-part hurdle model
%  combining a binary model for zero vs.\ positive precipitation and a Gaussian model,
%  truncated at zero, for the positive precipitation observations. The latter could
%  potentially better discriminate determinants of the occurrence of precipitation
%  vs.\ the amount of precipitation given that it occurs.
%}
%\end{supplement}

%\begin{supplement}[id=supp2]
%  \sname{Supplement 2}
%  \stitle{Stationwise Evaluation}
%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
%  \sdatatype{.pdf}" 
%  \sdescription{
%  In the case study presented in the main manuscript ``Distributional Regression
%  Forests for Probabilistic Precipitation Forecasting in Complex Terrain''
%  the novel distributional regression forests are evaluated and compared with
%  other distributional regression approaches on 95 meteorological stations.
%  Moreover, further details -- including a cross-validation analysis -- are
%  preformed for only one station, Axams. To show that Axams is a fairly typical
%  station and similar insights can be obtained for other stations as well,
%  this supplements presents the same analysis as in Section~3.3 of the main
%  manuscript for 14~further meteorological stations.
%}
%\end{supplement}

@@

This permutation test based tree algorithm is presented in 
\cite{Hothorn+Hornik+Zeileis:2006} as the CTree algorithm. 
A different framework to build a tree is provided by the MOB algorithm 
which is based on M-fluctuation tests (\citealt{Zeileis+Hothorn+Hornik:2008}).

@@

%As an alternative to QQ~plots based on randomized quantile residuals,
%probability integral transform (PIT, \citealt{Gneiting+Balabdaoui+Raftery:2007}) 
%histograms are a commonly used tool to assess how well-calibrated distributional 
%fits are. Figure~\ref{fig:axams-pit} shows out-of-sample PIT histograms for the same
%probabilistic forecasts as Figure~\ref{fig:axams-qq}, confirming that despite
%some deviations the overall calibration of all models is acceptable.

>>>>

\begin{itemize}
  \item \emph{Generalized linear models} (GLMs, \citealp{Nelder+Wedderburn:1972})
    encompassing an additional nonlinear link function for the conditional mean.
  \item \emph{Generalized additive models} (GAMs, \citealp{Hastie+Tibshirani:1986})
    allowing for smooth nonlinear effects in the covariates
    (Figure~\ref{fig:gam}, middle).
  \item \emph{Generalized additive models for location, scale, and shape}
    (GAMLSS, \citealp{Rigby+Stasinopoulos:2005}) adopting a probabilistic modeling
    approach. In GAMLSS, each parameter of a statistical distribution can depend
    on an additive predictor of the covariates comprising linear and/or
    smooth nonlinear terms (Figure~\ref{fig:gam}, right).
\end{itemize}
Thus, the above-mentioned models provide a broad toolbox for capturing different
aspects of the response (mean only vs.\ full distribution) and different types
of dependencies on the covariates (linear vs.\ nonlinear additive terms).

@@

In summary, the models discussed so far provide a broad and powerful toolset for
parametric distributional fits depending on a specified set of additive linear or
smooth nonlinear terms. A rather different approach to capturing the dependence on
covariates are tree-based models.
\begin{itemize}
  \item \emph{Regression trees} (\citealp{Breiman+Friedman+Stone:1984}) recursively
    split the data into more homogeneous subgroups and can thus capture abrupt shifts
    (Figure~\ref{fig:tree}, top left) and approximate nonlinear functions.
    Furthermore, trees automatically carry out a forward selection of covariates and
    their interactions. 
  \item \emph{Random forests} (\citealp{Breiman:2001}) average the predictions of
    an ensemble of trees fitted to resampled versions of the learning data. This
    stabilizes the recursive partitions from individual trees and hence better
    approximates smooth functions (Figure~\ref{fig:tree}, top middle).
\end{itemize}
While classical regression trees and random forests only model the mean of the response
we propose to follow the ideas from GAMLSS modeling -- as outlined
in Figure~\ref{fig:gam} -- and combine tree-based methods with parametric distributional models,
yielding two novel techniques:
\begin{itemize}
  \item \emph{Distributional regression trees} (for short: \emph{distributional trees}) split the data into more homogeneous groups with
    respect to a parametric distribution, thus capturing changes in any distribution
    parameter like location, scale, or shape (Figure~\ref{fig:tree}, bottom left).
  \item \emph{Distributional regression forests} (for short: \emph{distributional forests}) utilize an ensemble of distributional trees
    for obtaining stabilized and smoothed parametric predictions (Figure~\ref{fig:tree},
    top right).
\end{itemize}

@@

Learning and validation data consist of observed 
daily precipitation sums provided by the National Hydrographical Service 
(\citealp{ehyd}) and numerical weather forecasts from the U.S.~National
Oceanic and Atmospheric Administration (NOAA).
Both, observations and forecasts are available for 1985--2012 and
the analysis is exemplified using July, the month with the most precipitation
in Tyrol.

@@

The predictive performance in terms of full probabilistic forecasts is
assessed using the continuous ranked probability score (CRPS, \citealp{Hersbach:2000}).
For each of the models this assesses the discrepancy of the
predicted distribution function $F$ from the observation $y$ by
$$
\text{CRPS}(y, F) = \int_{-\infty}^\infty (F(z) - \mathbf{1}(y \leq z))^2 dz
$$
where \(\mathbf{1}(\cdot)\) is the indicator function. In the subsequent
applications, the mean CRPS is always evaluated out of sample, 
either using cross-validation or a hold-out data set (2009--2012)
that was not used for learning (1985--2008). CRPS is a proper scoring rule
\citep{Gneiting+Raftery:2007} often used within the meteorological community.
Lower values indicate better performance.

@@

To assess calibration Figure~\ref{fig:axams-qq} shows residual QQ~plots for
out-of-sample predictions (2009--2012) from the different models trained on
1985--2008. Due to the point masses at zero 100 draws from the randomized
quantile residuals \citep{Dunn+Smyth:1996} are plotted in semi-transparent
gray. Overall, the randomized quantile residuals conform quite well with the
theoretical standard normal quantile (i.e., form a straight line close to the 
diagonal), indicating that all four models are sufficiently well calibrated. 
\hlc{This is also supported by the corresponding probability integral transform}
(PIT, \citealp{Gneiting+Balabdaoui+Raftery:2007}) \hlc{histograms for station
Axams in Supplement B} \citep{Schlosser+Hothorn+Stauffer:2018b} \hlc{which 
contains a more detailed explanation of residual QQ plots and PIT histograms.}

@@

\subsubsection{Distributional specifications}
\hlc{Choosing an adequate distributional family is a crucial step and forms
the basis required to build a well fitting model. Therefore, this
decision should be made thoughtfully. 
In the presented application we applied a Gaussian distribution left-censored 
at zero as this has proven to be a well working choice in
earlier literature} \citep{Stauffer+Mayr+Messner:2017}. 
\hlc{To test for robustness against distributional misspecifation 
different distributional assumptions have been considered as well. 
In Supplement A} \citep{Schlosser+Hothorn+Stauffer:2018a} \hlc{the exact
same evaluations as in} Section~\ref{sec:all} \hlc{are carried out for 15
observation stations, once employing a Gaussian hurdle model and once
with a logistic distribution left-censored at zero in order to account
for heavier tails. Under both assumptions the obtained results 
are similar to those presented in the paper applying a Gaussian distribution
left-censored at zero. In some settings the hurdle model leads to slight 
improvements, but this comes with more variability (especially for EMOS and the 
boosted GAMLSS). For the boosted GAMLSS the left-censored logistic distribution
shows slightly better results regarding CRPS skill score, however, only within 
a range of less than $2\%$ and not for all stations. 
Altogether, the distributional forest has proven to be 
very robust against distributional misspecifications.}

@@

\subsubsection*{East Tyrol}
Figure~\ref{fig:all-map} \hlc{shows that, contrary to the rest of Tyrol, 
for the majority of stations in East Tyrol the prespecified GAMLSS performs
best in this particular setting (learning data: 1985--2008, testing data:
2009--2012). East Tyrol is located south of the main Alpine Ridge and 
exhibits different climatological characteristics than the part north of 
the Alpine Ridge. It seems that the choice of covariates and their interactions 
we made for the prespecified GAMLSS method is particularly adequate for the 
region of East Tyrol for this specific analysis. However, the advantages of 
the prespecified GAMLSS models in this particular setting (training: 1985--2008, 
validation: 2009--2012) even out in a cross-validation framework for most of 
the stations in East Tyrol. These results are presented in Supplement B} 
\citep{Schlosser+Hothorn+Stauffer:2018b} \hlc{where the exact same evaluations as in} 
Section~\ref{sec:axams} \hlc{for station Axams have been carried out 
for 14 other stations.}

\subsubsection*{Choice of station Axams}
%% choice of station Axams mit Verweis auf Supplement stationwise
\hlc{Axams was chosen for the more extensive evaluations presented
in this paper as it is the observation station being closest 
to Innsbruck, the capital of Tyrol and the work place of 
three of the authors, but not based on any other particular 
features. Furthermore, it has been checked, whether it 
is an adequate representative of all stations regarding 
the results of our evaluations. This is also confirmed by 
the extended analysis in Supplement B} \citep{Schlosser+Hothorn+Stauffer:2018b}
\hlc{where the 15 considered stations also include some of those where 
the forest does not perform better than the other models in the setting 
presented in} Figure~\ref{fig:all-map}. \hlc{It shows that the differences in 
performance even out in the cross-validation framework where the forest 
performs at least equally well compared to the best of the other models 
at most stations.}

@@

The proposed methods are in the \textsf{R} package \textbf{disttree}
(version~\Sexpr{packageVersion("disttree")}) based on the \textbf{partykit}
package (version~\Sexpr{packageVersion("partykit")}), both available 
on \textsf{R}-Forge at
(\url{https://R-Forge.R-project.org/projects/partykit/}). The function
\code{distfit} fits distributional models by maximum likelihood, which is
used as the basis for the tree-building function \code{disttree}, upon which
the \code{distforest} is built. All functions can either be used with
GAMLSS family objects from the \textsf{R} package \textbf{gamlss.dist} 
(\citealp{Stasinopoulos+Rigby:2007}, version~\Sexpr{packageVersion("gamlss.dist")})
or with custom lists containing all required information about the distribution family.

@@

The fitted distributional forest for July 24 and
observation station Axams (including Figure~\ref{fig:axams-dist})
is reproducible using 
\code{demo("RainAxams", package = "disttree")}.
This also includes fitting the other zero-censored
Gaussian models considered in this paper and generating the corresponding 
QQ~plots (Figure~\ref{fig:axams-qq}) and PIT histograms 
\citep[Supplement B]{Schlosser+Hothorn+Stauffer:2018b}.
Full replication of all results can be obtained with
\code{demo("RainTyrol", package = "disttree")}
requiring the companion \textsf{R} package \textbf{RainTyrol} (version~\Sexpr{packageVersion("RainTyrol")}),
also available within the \textsf{R}-Forge~project.
\hlc{The results presented in Supplement A} \citep{Schlosser+Hothorn+Stauffer:2018a} and 
Supplement B \citep{Schlosser+Hothorn+Stauffer:2018b} \hlc{can be reproduced using}
\code{demo("RainDistributions", package = "disttree")} and 
\code{demo("RainStationwise", package = "disttree")} respectively.

@@

%\begin{supplement}[id=supp1]
%  \sname{Supplement A}
%  \stitle{Different Response Distributions}
%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
%  \sdatatype{.pdf}" 
%  \sdescription{
%  In the case study presented in the main manuscript ``Distributional Regression
%  Forests for Probabilistic Precipitation Forecasting in Complex Terrain''
%  the novel distributional regression forests are based on a Gaussian response distribution,
%  left-censored at zero to accomodate both the non-negativity of precipitation and
%  the point mass at zero. To assess the goodness of fit of this response distribution
%  this supplement employs the same evaluations as in the main manuscript but based on
%  two other distributional assumptions: A logistic distribution, left-censored at zero,
%  is employed to potentially better capture heavy tails -- and a two-part hurdle model
%  combining a binary model for zero vs.\ positive precipitation and a Gaussian model,
%  truncated at zero, for the positive precipitation observations. The latter could
%  potentially better discriminate determinants of the occurrence of precipitation
%  vs.\ the amount of precipitation given that it occurs.
%}
%\end{supplement}

%\begin{supplement}[id=supp2]
%  \sname{Supplement B}
%  \stitle{Stationwise Evaluation}
%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
%  \sdatatype{.pdf}" 
%  \sdescription{
%  In the case study presented in the main manuscript ``Distributional Regression
%  Forests for Probabilistic Precipitation Forecasting in Complex Terrain''
%  the novel distributional regression forests are evaluated and compared with
%  other distributional regression approaches on 95 meteorological stations.
%  Moreover, further details -- including a cross-validation analysis -- are
%  preformed for only one station, Axams. To show that Axams is a fairly typical
%  station and similar insights can be obtained for other stations as well,
%  this supplements presents the same analysis as in Section~3.3 of the main
%  manuscript for 14~further meteorological stations.
%}
%\end{supplement}

@@

This permutation test based tree algorithm is presented in 
\cite{Hothorn+Hornik+Zeileis:2006} as the CTree algorithm. 
A different framework to build a tree is provided by the MOB algorithm 
which is based on M-fluctuation tests (\citealp{Zeileis+Hothorn+Hornik:2008}).

@@

%As an alternative to QQ~plots based on randomized quantile residuals,
%probability integral transform (PIT, \citealp{Gneiting+Balabdaoui+Raftery:2007}) 
%histograms are a commonly used tool to assess how well-calibrated distributional 
%fits are. Figure~\ref{fig:axams-pit} shows out-of-sample PIT histograms for the same
%probabilistic forecasts as Figure~\ref{fig:axams-qq}, confirming that despite
%some deviations the overall calibration of all models is acceptable.

### 2018-10-31 pkg/disttree/inst/draft/rejoinder.Rnw
changed numbering of supplements (A and B instead of 1 and 2)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2369 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 9093ae989a286d620c85bcfcb92b68ad87d3c504

\item \emph{Supplement 1: Different response distributions.}\\
To assess the appropriateness and robustness of the distributional
assumption (censored Gaussian) made in the main manuscript, further
response distributions (two-part Gaussian hurdle; censored
logistic) are evaluated to potentially better capture lower
and upper tails, respectively. The supplement follows the
evaluations performed for all 95 observation stations 
in the main manuscript using all three response distributions
for a subset of 15 stations.

\item \emph{Supplement 2: Stationwise evaluation.}\\
To show that the more extensive cross-validation results -- obtained only
for one station (Axams) in the main manuscript -- are fairly typical,
they are extended to 14 other observation stations. This also incorporates
results that were previously presented in Appendix B which has consequently
been omitted from the main manuscript.

@@

% \vspace*{-2.8cm}
% \hspace*{-0.3cm} %\includegraphics{zeileis-signature}

@@

We have measured fitting time and prediction time for our 
application. For observation-by-observation predictions -- as
done in real-time -- the computational cost for the distributional
forest is comparable to the GAMLSS approaches. The former have
to re-fit the model using the observation-specific weights
while the latter has to set up the model matrices for all the
additive terms. However, when predicting a set of observations
the costs for the GAMLSS are lower (due to leveraging matrix
operations) and for the distributional forest higher (as it
still operates observation-wise) but not prohibitively so.
The exact results and times can be found in Supplement 2
(Stationwise Evaluation).

@@

Axams was chosen as the observation station being closest 
to Innsbruck, the capital of Tyrol and the work place of 
three of the authors, but not based on any other particular 
features. Furthermore, it has been checked whether it 
is an adequate representative of all stations regarding 
the results of our evaluations. This is now also 
confirmed by further analysis in Supplement 2 (Stationwise Evaluation) 
and discussed in Section~4.

@@

The residual QQ plots (Dunn et al.\ 1996) are built on top of
the probability integral transform (PIT, Gneiting et al.\ 2007).
Therefore they are invariant to transformations that preserve
the PIT, as is the case here. A more detailed 
explanation can be found in Supplement 2 (Stationwise Evaluation).

@@

The advantages of the GAMLSS model even out at most of the stations when looking 
on the results of a cross-validation. We have extended Section~4 to discuss this 
issue. The results of an extended evaluation including a cross-validation for a 
set of $15$ stations can be found in Supplement 2 (Stationwise Evaluation).

@@

Yes, in distributional forests the parameter vector has to be estimated 
separately for each new prediction as it is based on
observation-specific weights. For that reason, making 
predictions for \emph{a set of} new observations is more 
time consuming for the forest (where predictions are obtained
separately) than the GAMLSS models (where matrix operations
can be leveraged). However, when making predictions observation
by observation -- as done in real-time -- the time required is
comparable for the forest and GAMLSS. The exact results 
regarding fitting and prediction times can now be found in
Supplement 2 (Stationwise Evaluation).

@@

The distributional forest selects very similar covariates as included in the
prespecified GAMLSS model. For both models, total precipitation (\texttt{tp})
and total column-integrated condensate (\texttt{tcolc}) are the two most important
covariates. Due to the randomization within the nodes, the distributional forest selects different variations of the same covariate (e.g., \texttt{tp\_max}, \texttt{tp\_mean},
\texttt{tp\_min}; see Figure~6 in the main manuscript and Supplement 2) while
the prespecified GAMLSS mainly only uses one specific variation (\texttt{tp\_max}).
The same is true for variations of \texttt{tcolc\_mean\_*}, \texttt{tcolc\_sprd\_*}
and \texttt{tcolc\_sprd\_*}.

@@

We have carried out the same evaluations as presented in the main manuscript and 
present the full results in Supplement 1 (Different Response Distributions). 
The main insights from this are also discussed in Section~4:
At most stations the censored Gaussian distribution from the main
manuscript performs at least as well as the two alternative
distributions. %, if not better. 
Moreover, the distributional forest proves to be rather robust against distributional
misspecifications. Therefore, the discussion in the main manuscript is still based
on the censored Gaussian distribution.

@@

In order to investigate these questions we have carried out the full
analysis as presented for station Axams in the main manuscript for 14
further observation stations. It shows that at most stations the 
distributional forest is competitive with the best among the other 
methods or even better. The exact results are provided in 
Supplement 2 (Stationwise Evaluation) and discussed in Section~4. 
Especially for meteorologists intending to apply model output 
statistics (MOS) for dozens or hundreds of observation stations 
(e.g., Wahl 2014, Rasp and Lerch 2018)
the distributional forest provides a good and robust alternative. 
Computational costs are clearly higher than for a basic EMOS model, 
but similar to a prespecified GAMLSS and lower than for GAMLSS
using variable selection by cross-validation. Fitting and forecasting
times are reported now in Supplement~2 (Stationwise Evaluation).

@@

Thank you for your advice. We have kept only the details regarding
the presented methodology in the appendix of the main manuscript
(Appendix A) and extended Appendix B to an online supplement
(Supplement 2: Stationwise Evaluation).

@@

The PIT histograms and the residual QQ plots visualize the same
results on different scales. The out-of-sample PIT values are 
either plotted in a histogram or plotted against the corresponding
quantiles of the standard normal distribution. A more detailed 
explanation can be found in Supplement 2 (Stationwise Evaluation) 
which contains both, PIT histograms and residual QQ plots, 
for 15 stations. In the manuscript we decided to only show the 
residual QQ plot as it allows for a better analysis of the 
behavior in the upper tail compared to the PIT histogram.

>>>>

\item \emph{Supplement A: Different response distributions.}\\
To assess the appropriateness and robustness of the distributional
assumption (censored Gaussian) made in the main manuscript, further
response distributions (two-part Gaussian hurdle; censored
logistic) are evaluated to potentially better capture lower
and upper tails, respectively. The supplement follows the
evaluations performed for all 95 observation stations 
in the main manuscript using all three response distributions
for a subset of 15 stations.

\item \emph{Supplement B: Stationwise evaluation.}\\
To show that the more extensive cross-validation results -- obtained only
for one station (Axams) in the main manuscript -- are fairly typical,
they are extended to 14 other observation stations. This also incorporates
results that were previously presented in Appendix B which has consequently
been omitted from the main manuscript.

@@

\vspace*{-2.8cm}
\hspace*{-0.3cm} 
\includegraphics[width = 0.23\textwidth]{signature-Schlosser}

@@

We have measured fitting time and prediction time for our 
application. For observation-by-observation predictions -- as
done in real-time -- the computational cost for the distributional
forest is comparable to the GAMLSS approaches. The former have
to re-fit the model using the observation-specific weights
while the latter has to set up the model matrices for all the
additive terms. However, when predicting a set of observations
the costs for the GAMLSS are lower (due to leveraging matrix
operations) and for the distributional forest higher (as it
still operates observation-wise) but not prohibitively so.
The exact results and times can be found in Supplement B
(Stationwise Evaluation).

@@

Axams was chosen as the observation station being closest 
to Innsbruck, the capital of Tyrol and the work place of 
three of the authors, but not based on any other particular 
features. Furthermore, it has been checked whether it 
is an adequate representative of all stations regarding 
the results of our evaluations. This is now also 
confirmed by further analysis in Supplement B (Stationwise Evaluation) 
and discussed in Section~4.

@@

The residual QQ plots (Dunn et al.\ 1996) are built on top of
the probability integral transform (PIT, Gneiting et al.\ 2007).
Therefore they are invariant to transformations that preserve
the PIT, as is the case here. A more detailed 
explanation can be found in Supplement B (Stationwise Evaluation).

@@

The advantages of the GAMLSS model even out at most of the stations when looking 
on the results of a cross-validation. We have extended Section~4 to discuss this 
issue. The results of an extended evaluation including a cross-validation for a 
set of $15$ stations can be found in Supplement B (Stationwise Evaluation).

@@

Yes, in distributional forests the parameter vector has to be estimated 
separately for each new prediction as it is based on
observation-specific weights. For that reason, making 
predictions for \emph{a set of} new observations is more 
time consuming for the forest (where predictions are obtained
separately) than the GAMLSS models (where matrix operations
can be leveraged). However, when making predictions observation
by observation -- as done in real-time -- the time required is
comparable for the forest and GAMLSS. The exact results 
regarding fitting and prediction times can now be found in
Supplement B (Stationwise Evaluation).

@@

The distributional forest selects very similar covariates as included in the
prespecified GAMLSS model. For both models, total precipitation (\texttt{tp})
and total column-integrated condensate (\texttt{tcolc}) are the two most important
covariates. Due to the randomization within the nodes, the distributional forest selects different variations of the same covariate (e.g., \texttt{tp\_max}, \texttt{tp\_mean},
\texttt{tp\_min}; see Figure~6 in the main manuscript and Supplement B) while
the prespecified GAMLSS mainly only uses one specific variation (\texttt{tp\_max}).
The same is true for variations of \texttt{tcolc\_mean\_*}, \texttt{tcolc\_sprd\_*}
and \texttt{tcolc\_sprd\_*}.

@@

We have carried out the same evaluations as presented in the main manuscript and 
present the full results in Supplement A (Different Response Distributions). 
The main insights from this are also discussed in Section~4:
At most stations the censored Gaussian distribution from the main
manuscript performs at least as well as the two alternative
distributions. %, if not better. 
Moreover, the distributional forest proves to be rather robust against distributional
misspecifications. Therefore, the discussion in the main manuscript is still based
on the censored Gaussian distribution.

@@

In order to investigate these questions we have carried out the full
analysis as presented for station Axams in the main manuscript for 14
further observation stations. It shows that at most stations the 
distributional forest is competitive with the best among the other 
methods or even better. The exact results are provided in 
Supplement B (Stationwise Evaluation) and discussed in Section~4. 
Especially for meteorologists intending to apply model output 
statistics (MOS) for dozens or hundreds of observation stations 
(e.g., Wahl 2014, Rasp and Lerch 2018)
the distributional forest provides a good and robust alternative. 
Computational costs are clearly higher than for a basic EMOS model, 
but similar to a prespecified GAMLSS and lower than for GAMLSS
using variable selection by cross-validation. Fitting and forecasting
times are reported now in Supplement~2 (Stationwise Evaluation).

@@

Thank you for your advice. We have kept only the details regarding
the presented methodology in the appendix of the main manuscript
(Appendix A) and extended Appendix B to an online supplement
(Supplement B: Stationwise Evaluation).

@@

The PIT histograms and the residual QQ plots visualize the same
results on different scales. The out-of-sample PIT values are 
either plotted in a histogram or plotted against the corresponding
quantiles of the standard normal distribution. A more detailed 
explanation can be found in Supplement B (Stationwise Evaluation) 
which contains both, PIT histograms and residual QQ plots, 
for 15 stations. In the manuscript we decided to only show the 
residual QQ plot as it allows for a better analysis of the 
behavior in the upper tail compared to the PIT histogram.
